When two-point correlation function fails to describe a structure...

### 2024-03-07 paper.tex
Describe why we do a substitution t^{-1} = z + z^{-1}

\begin{definition}
  A Z-transform of a sequence $\{a_0, a_1, \dots, a_n\}$ is a polynomial
  $\sum_{i=0}^n a_i z^{-i}$.
\end{definition}
Z-transform is much like the discrete Fourier transform, particulary it has the
same convolution theorem:
\begin{theorem}
  If $g(z)$ and $h(z)$ are images of sequences $a$ and $b$ under Z-transform,
  then $g(z)h(z) = f(z)$ where $f(z)$ is an image of $a*b$ under Z-transform
  and $a*b$ is a discrete convolution of $a$ and $b$:
  \begin{equation*}
    (a*b)_k = \sum_{i=0}^{n} a_i b_{k-i}
  \end{equation*}
  \label{th:conv}
\end{theorem}
In the case of one-dimensional two-phase porous media $a$ the definition
\ref{eq:s2} can be rewritten as follows
\begin{equation*}
  S_2(r) = \frac{\sum_{i}a_ia_{i+r}}{N(r)}
\end{equation*}
where $N(r)$ is a normalization factor which does not depend on $a$ and can be
omitted for simplicity. The numerator is a discrete autocorrelation which is
equal to a convolution of $a$ with ``reversed'' $a$, i.e. $i$-th element of $a$
becomes $-i$-th element of a reversed sequence. Let $f(z)$ be an image of $a$
under Z-transform. Then an image $s_2(z)$ of the sum above can be calculated
using \cref{th:conv}.
\begin{equation}
  s_2(z) = f(z)f(z^{-1})
  \label{eq:s2z}
\end{equation}
Making a substitution $t^{-1} = z + z^{-1}$ and inverting Z-transform for
$s_2(t)$ we will get a sequence $S_2(r) = (S_2(0), S_2(1), S_2(2), \dots)$ up to
a normalization by $N(r)$. The meaning of this sequence is a number of positive
elements in $a$ distanced by $r$.
\begin{example}
  Let $a$ be a sequence $(1, 0, 1, 0, 0, 1)$. Its Z-transform is
  $f(z) = 1 + z^{-2} + z^{-5}$.
  \begin{align*}
    s_2(z) &= f(z)f(z^{-1}) \\
    &= (1 + z^{-2} + z^{-5})(1 + z^2 + z^5) \\
    &= (1 + z^2 + z^5) + (z^{-2} + 1 + z^3) + (z^{-5} + z^{-3} + 1) \\
    &= 3 + (z^2 + z^{-2}) + (z^3 + z^{-3}) + (z^5 + z^{-5})
  \end{align*}
  Thus, we have $s(t) = 3 + t^{-2} + t^{-3} + t^{-5}$. Inverting Z-transform, we
  obtain a sequence $(3, 0, 1, 1, 0, 1)$, which means that $a$ has three
  elements in phase $1$, zero adjacent elements in phase $1$, one pair which is
  distanced by 2 (elements with indices 0 and 2), one pair distanced by 3
  (indices 2 and 5) and one pair distanced by 5 (indices 0 and 5).
\end{example}

>>>>

\begin{definition}
  A Z-transform of a sequence $\{a_0, a_1, \dots, a_n\}$ is a polynomial
  $\sum_{i=0}^n a_i z^{-i}$.
\end{definition}
Z-transform is much like the discrete Fourier transform, particulary it has the
same convolution theorem:
\begin{theorem}
  If $g(z)$ and $h(z)$ are images of sequences $a$ and $b$ under Z-transform,
  then $g(z)h(z) = f(z)$ where $f(z)$ is an image of $a*b$ under Z-transform
  and $a*b$ is a discrete convolution of $a$ and $b$:
  \begin{equation*}
    (a*b)_k = \sum_{i=0}^{n} a_i b_{k-i}
  \end{equation*}
  \label{th:conv}
\end{theorem}
In the case of one-dimensional two-phase porous media $a$ the definition
\ref{eq:s2} can be rewritten as follows
\begin{equation*}
  S_2(r) = \frac{\sum_{i}a_ia_{i+r}}{N(r)}
\end{equation*}
where $N(r)$ is a normalization factor which does not depend on $a$ and can be
omitted for simplicity. The numerator is a discrete autocorrelation which is
equal to a convolution of $a$ with ``reversed'' $a$, i.e. $i$-th element of $a$
becomes $-i$-th element of a reversed sequence. Let $f(z)$ be an image of $a$
under Z-transform. Then an image $s_2(z)$ of the sum above can be calculated
using \cref{th:conv}.
\begin{equation}
  s_2(z) = f(z)f(z^{-1})
  \label{eq:s2z}
\end{equation}
Because $S_2(r) = S_2(-r)$ it is convenient to make a substitution $t^{-1} = z +
z^{-1}$ and consider a function of a non-negative argument. Inverting
Z-transform for $s_2(t)$ we will get a sequence
$S_2(r) = (S_2(0), S_2(1), S_2(2), \dots)$ up to a normalization by $N(r)$. The
meaning of this sequence is a number of positive elements in $a$ distanced by
$r$.
\begin{example}
  Let $a$ be a sequence $(1, 0, 1, 0, 0, 1)$. Its Z-transform is
  $f(z) = 1 + z^{-2} + z^{-5}$.
  \begin{align*}
    s_2(z) &= f(z)f(z^{-1}) \\
    &= (1 + z^{-2} + z^{-5})(1 + z^2 + z^5) \\
    &= (1 + z^2 + z^5) + (z^{-2} + 1 + z^3) + (z^{-5} + z^{-3} + 1) \\
    &= 3 + (z^2 + z^{-2}) + (z^3 + z^{-3}) + (z^5 + z^{-5})
  \end{align*}
  Thus, we have $s(t) = 3 + t^{-2} + t^{-3} + t^{-5}$. Inverting Z-transform, we
  obtain a sequence $(3, 0, 1, 1, 0, 1)$, which means that $a$ has three
  elements in phase $1$, zero adjacent elements in phase $1$, one pair which is
  distanced by 2 (elements with indices 0 and 2), one pair distanced by 3
  (indices 2 and 5) and one pair distanced by 5 (indices 0 and 5).
\end{example}
